\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{SmarteRun: Coding The Portable Coach}

### 2014-05-13 final/article.tex
Added combining lines figure
rewrote abstract a bit

\noindent The pairing of physical activity and music is a natural, common occurrence in modern society.  Today, there are many smartphone and smart device applications focused on this pairing, specifically catering towards runners. Currently, these applications make no attempt to combine music and running in a novel way, instead allowing only for both activities to occur simultaneously. As the prevalence of smart devices continues to alter the technological landscape, the opportunity for the union of this pair is becoming more feasible.  We consider psychological studies that consider the rhythmic nature of human exercise and the effects of music upon this rhythm  We then synthesize a model by which previous activity can predict future run performance and capabilities. We then analyze these predictions and discuss further improvements which can be made to our models. We conclude with the application of this predictive model to music playback and the ability to cater music at speeds which accentuate runners' performance and create the running coach athletes didn't know they had.

@@

	All in all, it is clear that music is effective in influencing the performance of runners, providing value to benchmark performance, enjoyment of the workout, and regularity of keeping to a regimen. (Karageorghis)\\

	Other attempts at using past exercise data to predict future performance are scarce, however there is one major example. In 2013, the Boston Marathon had to be shutdown before many participants could finish due to an act of terrorism. In order to determine run times for the unfinished runners, a group from the Harvard School of Public Health was employed to predict performance for the incomplete times. The investigation determined the best approach was to use a K-nearest-neighbor algorithm searching for similar other runs in a dataset of past run data, and then combining them to extrapolate to the rest of the run. This method far exceeded the "constant pace" assumption others had proposed, as well as several other statistical attempts such as the ANOVA method. This method worked very well as it predicted 98\% of runs to within 10 minutes. While this is a large error for normal runs, 10 minutes is small on marathon scale. For 80\% of runners, that error is cut in half. (Hammerling)\\

@@

	Our prediction algorithm implements a merging of our closest neighbors, iterating through each time interval for which an average of the elapsed distances is produced using geometric averaging of coordinates.  We achieve reasonable accuracy in this process by using an exponential decay weighting, by which runs further in distance from our desired prediction are given exponentially lower weighting.  By this we have our weighted output value:. We defined our gamma value such that we give a neighboring run <1\% weighting after a 5km distance from our desired output.  We believe this difference is the value at which we are considering distinctly separate types of runs.  We note that generally for every given kilometer away from our desired output, the weighting of a neighbor is essentially cut in a third.\\

	Using this predicted graph R(x,y,z) we provide tangent-line and secant-line predictions such that we provide a recommended pace for a runner such that given their current performance they achieve, exceed, or relax compared to their usual performance.  This step essentially merges our predictive data with the runner'۪s current activity and would be the useful user-facing data provided in a retail application of our algorithms.\\

@@

Here we present the full approach for measuring accuracy:

>>>>

\noindent Pairing physical activity and music is commonplace in today's society, particularly in the running community.  Many smart devices or their applications are able to play music and track physical performance; some can even do both simultaneously. However, this technology also presents an opportunity for growth. Here we propose a synthesis of music playing and performance tracking to increase a runner's potential. We justify our proposal by considering psychological studies of the rhythmic nature of human exercise and the effects of music upon this rhythm. The model we propose uses the K-nearest-neighbor algorithm on previous activity to predict future run performance and potential. Analyzing these predictions suggests that this model can optimize a runner's performance under particular conditions and types of runs. Improvements which can be made to our model include the ability to match on more more of the data available about a current run. Applying this predictive model to music playback and speed can create the running coach athletes did not know they had.

@@

	All in all, it is clear that music is effective in influencing the performance of runners, providing value to benchmark performance, enjoyment of the workout, and regularity of keeping to a regimen. (Karageorghis, et al.)\\

	Other attempts at using past exercise data to predict future performance are scarce, however there is one major example. In 2013, the Boston Marathon had to be shutdown before many participants could finish due to an act of terrorism. In order to determine run times for the unfinished runners, a group from the Harvard School of Public Health was employed to predict performance for the incomplete times. The investigation determined the best approach was to use a K-nearest-neighbor algorithm searching for similar other runs in a dataset of past run data, and then combining them to extrapolate to the rest of the run. This method far exceeded the "constant pace" assumption others had proposed, as well as several other statistical attempts such as the ANOVA method. This method worked very well as it predicted 98\% of runs to within 10 minutes. While this is a large error for normal runs, 10 minutes is small on marathon scale. For 80\% of runners, that error is cut in half. (Hammerling, et al.)\\

@@

	Our prediction algorithm implements a merging of our closest neighbors, iterating through each time interval for which an average of the elapsed distances is produced using geometric averaging of coordinates.  We achieve reasonable accuracy in this process by using an exponential decay weighting, by which runs further in distance from our desired prediction are given exponentially lower weighting.  By this we have our weighted output value: $a_k = \delta^{\mid z-z_k\mid}$. We defined our delta value such that we give a neighboring run <1\% weighting after a 5km distance from our desired output.  We believe this difference is the value at which we are considering distinctly separate types of runs.  We note that generally for every given kilometer away from our desired output, the weighting of a neighbor is essentially cut in a third.\\

 \begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{combining.png}
\caption{A graph showing the 5 nearest neighbors (thin lines) averaged into the predicted run model (thick line). }
\end{center}
\end{figure}

	Using this predicted graph R(x,y,z) we provide tangent-line and secant-line predictions such that we provide a recommended pace for a runner such that given their current performance they achieve, exceed, or relax compared to their usual performance.  This step essentially merges our predictive data with the runner's current activity and would be the useful user-facing data provided in a retail application of our algorithms.\\

@@
